\chapter*{Abstract}

Alzheimer's disease is a devastating neurodegenerative disorder characterised by progressive intracellular accumulation of hyperphosphorylated tau and extracellular deposition of amyloid-beta (A$\beta$). It affects over 50 million people worldwide with numbers expecting to triple by 2050. Despite recent success in identifying candidate risk genes for AD, the mechanisms underpinning disease progression remain unknown and a challenge to research efforts worldwide. An increasing number of transcriptome profiling studies implicate altered transcriptional regulation and RNA splicing in the development of AD pathology. However, such studies are methodologically-flawed due to inherent limitations of standard RNA-sequencing approaches to capture full-length transcripts critical for transcriptome assembly.  

Consequently, the primary aim of this thesis was to use two long-read sequencing approaches, isoform sequencing (Pacific Biosciences) and nanopore cDNA sequencing (Oxford Nanopore Technologies), to assess progressive transcriptional and splicing variation associated with AD in a mouse model of tau pathology, rTg4510. By generating long reads that span full-length transcripts, our studies revealed widespread RNA isoform diversity with unprecedented detection of novel transcripts not existing in genome annotations. Targeted sequencing of 20 AD-risk genes showed robust expression changes at the transcript level associated with tau accumulation. Our analyses provide a systematic evaluation of transcript usage, even in the absence of gene expression alterations, and highlight the importance of alternative RNA splicing as a mechanism underpinning gene regulation in AD development. 

In summary, this thesis presents the first comprehensive study to leverage the power of long-read sequencing to assess AD transcriptional and splicing variation in a tau mouse model, with the capacity to dissect transcriptional alterations underlying AD pathology.       