\chapter*{Acknowledgements}

\begingroup
\singlespacing
While I had been warned of the difficulties of undertaking a PhD, I never would have imagined just how challenging and mentally taxing it could be. However, I have been blessed with the support of a number of people who have made this journey significantly easier and memorable. 

I would like to begin by thanking my supervisors – Professor Jonathan Mill, Dr Eilis Hannon and Dr Aaron Jeffries - for their invaluable guidance, patience and support throughout these four and a half years. Jon, thank you for the countless hours that you have spent patiently advising me throughout this PhD, and for always being so encouraging and enthusiastic about my results (even when I doubted them myself). Eilis, thank you for your insightful feedback and comments, and for being the person I could always turn to for a second, unbiased opinion. Aaron, thank you so much for sharing your invaluable expertise and knowledge be it in the lab or in bioinformatics, and for taking me under your wing; I cannot thank you enough for all those times that you have patiently reassured me in the lab when I thought I had messed up (again). Thank you to the three of you for continuing to inspire and show me what it takes to be a brilliant researcher. 

This work would not have been possible in it's current form without the funding from Eli-Lilly, and I would like to express my sincerest gratitude to Professor David Collier, Dr Emma Laing, Dr Zeshan Ahmed and Mr Gopuraja Dharmalingam for their ongoing interest and involvement. Your insightful comments and perspective (i.e."what question are we trying to answer here?") have shaped the development of this thesis and helped me to think more broadly. 

A special thanks to Dr Karen Moore and Paul O’Neill at the  Exeter Sequencing Service for sharing their expertise, particularly Karen for all of her guidance in optimising the Oxford Nanopore protocol.

I would like to thank all the members (past and present) of the Complex Disease Epigenetics group for being such a vibrant, enthusiastic and amazing group. Particularly to Thea and Gemma, I am so fortunate to have embarked on this journey with the both of you, to be on the same page and come back home to such a strong support network - thank you for making my day-to-day life at Exeter so much more enjoyable. To Isabel, without your dedication and passion, this PhD, which is partly an extension of your work, might not have existed, so thank you for all the times you have sat me down to discuss and brainstorm ideas. To Stef, thank you for all of your encouraging words both inside and outside of the lab, and for being exceptionally kind whenever I needed anything. 

A special thanks to Alan, who was always one step ahead of me on the PhD journey – thank you for listening to me rant (still) and resisting the urge to say “I told you so”. To all the members of the Belmont PhD group, especially Joshua - thank you for your wisdom, prayers and fellowship. To Lori, Tharini and Tommo - thank you for being the most wonderful friends and just being one phone call away no matter the distance. 

My sincere gratitude to my soon to be in-laws, Joe and Sandra, for making me feel so warmly welcomed at their home. Thank you for all the loving care that you have given me, for allowing me to visit almost every weekend, for the lovely meals, walks and spiritual conversations.
 
To my family – Ma, Han, Dad and Uncle Jabez – for their unconditional love, continuous support and unwavering patience, especially in the last few months of writing. Ma, thank you for always believing in me, for mentoring and guiding me; I would not be where I am without you. Han, thank you for picking up after me and shouldering responsibilities that should have been mine. Dad, thank you for never refusing to drive hours to drop my stuff off only to then pick them back up. Uncle Jabez, thank you for looking after the family without complaint and for taking me out of my research bubble - your resilience, passion and fortitude never ceases to amaze me.  

Finally, I would like to dedicate this thesis to my fiancé, Tim. Thank you for essentially being my fifth but uncredited supervisor. Despite having your own PhD, you listened to me go on repeatedly about alternative splicing, reviewed my statistical models and analyses, corrected my programming errors, opened my eyes to LaTeX, checked my grammar (even 5 minutes before I aimed to submit, though I still insist that "Lastly" should be a word), comforted me night and day, the list goes on... I would not have applied for a PhD without your unwavering encouragement, nor would I have finished it in time without your endless support. So thank you for everything, for putting up with me and committing to continually doing so :).

Last but foremost, praise and thanks be to God for keeping me motivated and sane throughout this PhD journey.

\endgroup